Active Object Search with a Mobile Device\\for People with Visual Impairments

### 2019-12-19 paper.tex
Fixing base repo to reflect camera ready

\title{Active Object Search with a Mobile Device for the Visually Impaired}

\author{
  \authorname{~}%Anon McAnon{\sup{1}}}
  \affiliation{~}%\sup{1}Anon University of Anon, Anonville, Anonistan}
  \email{~}%anon.anon@anon.com}
%\affiliation{\sup{1}Lincoln Centre for Autonomous Systems (L-CAS), University of Lincoln, Lincoln, UK}
%\email{jlock@lincoln.ac.uk}
}

\keywords{Active vision, object search, vision impairment, Markov Decision Process}

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments. The cameras embedded in these mobile devices, in particular, can be useful for many tasks, like reading signs or searching for objects in unknown environments. Although previous research in computer vision has looked at ways to solve these tasks by processing the video frames of the mobile camera, very little work has been done in {\em actively} guiding the user towards specific points of interest that maximise the effectiveness of the underlying visual algorithms. In this paper, we propose a control algorithm based on a Markov Decision Process (MDP) that generates real-time instructions to guide a user searching for a target object with a smartphone's camera. The solution is part of a more general active vision application for the visually impaired. An initial implementation of the system on an Android smartphone was experimentally evaluated with sighted participants to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help the visually impaired find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more normal and independent lives. Modern improvements in mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem. The work presented here is part of a project to assist the visually impaired (VI) navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017,bellotto2013}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}. Schematically, the actuators used in a typical control loop, shown in Figure~\ref{fig:control-loop}, are replaced by a human. %This project expands upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

\begin{figure}
  \centering
  \includegraphics[width=0.38\textwidth]{figures/system_use.png}
  \caption{The system in use during an experiment. }\label{fig:system-use}
\end{figure}

@@

\noindent Assistive vision and guidance for the VI is a growing research area. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.\ Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. These type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWiz system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since the VI person is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.

Previous researchers have implemented active search and perception strategies in robots and image classifiers~\cite{Bajcsy2017} in an attempt to optimise their classification and planning tasks, for example by exploiting the structured nature of human environments and object placements. 
%
Two research teams have recently implemented an active object search strategy into their image classifiers~\cite{caicedo2015active,gonzalez2015active}. Their approaches use different methods but conceptually similar models to generate windows of interest for visual classification. The size and locations of the windows within the image are generated using the spatial relationship between objects, taken from the SUNCG and PASCAL datasets~\cite{song2016ssc,Everingham10}, and are iteratively changed based on the output from the respective models. The advantage of their approaches is that fewer windows are generated and submitted to the classifier, resulting in lower object classification times, while still keeping state-of-the-art results for the accuracy.

@@

In summary, much research has been conducted on recognition of and guidance towards target objects, including active vision solutions for image classifiers and robotic systems. However, to our knowledge, no previous work has been done on {\em active} object search and guidance for humans, which would especially benefit VI users. In this paper, we implement such an active vision system with a human in the loop
% ~\cite{bellotto2013,lock2017portable}
that guides the user towards an out-of-view target object. Our system exploits prior knowledge of the objects spatial distribution within an indoor environment, learned from a dataset of real-world images, and the history of past object observations made during the search.
%With this, we hope to make the searching process less random and move towards a more natural and efficient object searching process to the benefit of the VI\@.

@@

\noindent The work presented in this paper is a fundamental step towards a more general project's goal to develop a stand-alone system that can guide a VI user to their destination with minimal user input or intervention. A complete system diagram is given is Figure~\ref{fig:control-loop}. This closed-loop system is conceptually similar to other classical control problems, where the difference between desired and actual state of a process is used to generate a control signal that changes the process itself.

In this case, the reference signal, $r$, is the object the user wishes to capture with the smartphone's camera. The goal of the control block, $K$, is to generate human interpretable instructions, $u$, to guide the user towards the target object. The process to be controlled involves a human, $H$, who interprets the instruction and executes a physical action, $u^*$, to actually manipulate the smartphone's camera, $P$. A new observation, $y$, with the latter is then fed back to the loop and the error signal, $e$, is updated accordingly.

Here we focus in particular on the implementation of the control module $K$. Two important points are considered in the design of the controller. Firstly, $K$ must be scenario-agnostic, meaning that objects could be placed in different places, unknown a priori. Secondly, the controller must be robust enough to handle incorrect interpretations by the human of the instructions $u$. This risk can be mitigated by the use of clear and simple instructions that helps $u$ track $u^*$ as closely as possible. The implementation of this controller is discussed in detail next.

@@

\noindent An MDP produces a policy of optimal actions for an agent to take in any given pre-defined state. In this case, the agent is defined as the guidance system and the policy is used to generate the next waypoint on the search path towards the target object. We assume the directly observable states and known state transitions probabilities. The MDP can be represented by the following 5-tuple:

\begin{equation}
  (\mathbf{S}, \mathbf{A}, \mathbf{T}, \mathbf{R}, \gamma)
\end{equation}
where $\mathbf{S}$ is a set of possible agent' states, $\mathbf{A}$ is a set of possible actions an agent can take in any given state, $\mathbf{T}$ is a set of state transition probabilities from state $s$ to state $s'$, with ${s,s'~\in~\mathbf{S}}$, and $\mathbf{R}$ is the reward the agent receives for reaching state $s'$ after executing action $a \in \mathbf{A}$ in state $s$. The scalar $\gamma$ is a discount factor which prioritises immediate over long-term rewards. Each of these elements are defined and discussed next.

@@

The policy produced by an MDP define the action an agent will take when it finds itself in any given state. In this case, the action is the direction of the next waypoint relatively to the current device's pose. The possible actions are four:

@@

To illustrate an example of actions sequence, let us consider the scene in Figure~\ref{fig:route-example}, which contains a number of simple, distinct objects (red boxes). The MDP guides the user in pointing the camera to the target object (mug at the bottom-left of the figure). It does this by inferring the current state, which depends on the object currently observed by the camera, and generating an action, or instruction, that leads the user to the target object.
%The states are sampled and the actions are generated in discretised steps as the user moves the camera's view in the instructed direction. %This discrete behaviour is similar to a waypoint navigation approach where the user goes from one waypoint to the next, until the user points the camera to target object. 
An action is considered completed, and therefore a new state reached, when the camera has rotated more than a predefined angle or a new object is detected.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/office_desk_example.png}
  \caption{An example action policies generated by the MDP to guide the user in pointing the camera from a random starting object (e.g. monitor) to a target object (e.g. mug). }\label{fig:route-example}
\end{figure}

@@

\noindent The reward $\mathbf{R}$ is the immediate reward that the agent receives after transitioning from state $s$ to state $s'$.
% and is given by 
% 
% \begin{equation}
%   R\in r(s, s').
% \end{equation} 
% 
The goal of the agent is to maximise its cumulative reward, therefore this is a very important parameter for producing a good action policy. In order to encourage the agent to find the target object as quick as possible, a relatively large positive reward should be assigned for successfully reaching the goal state, and a negative one in any other case. These parameters must be finely balanced to ensure an effective object search behaviour.

@@

\begin{equation}
  \begin{aligned}
    o\in \mathbf{O} = \{monitor, mouse, keyboard, window,\\mug, stationary, desk, nothing\}.
  \end{aligned}
\end{equation}

The spatial relationships between the objects in $\mathbf{O}$ are extracted from the OpenImage dataset~\cite{openimages}, which consists of 1.74M images containing 14.6M manually drawn and labelled bounding boxes around objects (see Figure~\ref{fig:openimage-example} for an example). The dataset is primarily aimed toward object recognition researchers to benchmark their models. In our case though, the bounding boxes and object labels are used to extract the spatial relationships between the different objects in $\mathbf{O}$. Since the camera perspectives and absolute distances between the objects in the images are not given, we can only extract the relationships in the basic action terms specified in $\mathbf{A}$, e.g. we can only say that object 1 is above object 2, but not how far above. Our relatively simple action-space is therefore suitable for the dataset's lack of information.

\begin{figure}
  \centering
  \begin{subfigure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/desk_example.png}
  \end{subfigure}
  \begin{subfigure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/mug_example.png}
  \end{subfigure}
  \caption{Examples of images from the OpenImage dataset containing objects from $\mathit{O}$~\cite{openimages}. }\label{fig:openimage-example}
\end{figure}

Figure~\ref{fig:obj-relationships} shows the spatial relationship between a subset of $\mathbf{O}$ (desk, keyboard and mouse). For example, when the agent is in state ${s=\langle o=mouse, n, v \rangle}$ and is searching for the object $o_{target} = keyboard$, there is a strong probability that the target object is on the mouse $LEFT$. The MDP of course will consider all of the objects' spatial relationships when generating the optimal policy.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/object_relationships.png}
  \caption{Examples of the spatial relationships between the desk, keyboard and mouse objects. Each square corresponds to the probability of executing an action (top square for $UP$, left square for $LEFT$, etc.)}\label{fig:obj-relationships}
\end{figure}

@@

\begin{equation}
  \alpha = \exp\Big(\frac{-i}{10~s_{total}}\Big) - 0.001,
\end{equation}
where $i$ is the episode index.  $\gamma$ is set to 0.95 to prioritise long-term reward.

Our MDP has a relatively small state-action space. Therefore, a solution can be found in a reasonable amount of time. However, it should be noted that adjusting the angle interval, or adding more actions or objects, can easily lead to an intractable size of the state-space.

\subsubsection{Android Implementation}\label{sec:system-implementation}

We incorporated the trained MDP-based controller into an app (see Figure~\ref{fig:system-screenshot}) for an Asus ZenPhone AR smartphone, running Android 7.0, with the Google's augmented reality toolkit (ARCore), which provides the necessary 3D pose of the device. No further software or hardware modifications were required. This app is responsible for generating the guidance instructions and tracking the camera sensor ($K$ and $P$ blocks in Figure~\ref{fig:control-loop}) throughout a search session. Tracking the camera's pose allows the app to infer the current state and choose the optimal action to take next.

\begin{figure}
  \centering
  \includegraphics[width=0.35\textwidth]{figures/system_screenshot2.png}
  \caption{A screenshot of the guidance interface showing a waypoint (based on an 3D Android model), the object scanner area and the guidance instruction arrows. }\label{fig:system-screenshot}
\end{figure}

The app uses a 6$\times$6 discretised and wrapped radial grid to simplify the tracking and waypoint generation processes. A policy action is converted by the app into a new search waypoint centred on a cell of the radial grid, e.g.\ an `$UP$' action will generate a waypoint one grid cell above the current the camera's orientation. Note that this cell is not part of the MDP' state: the radial grid is only used to discretise the pan-tilt movements of the camera and to guarantee a minimum angular variations between subsequent actions. The grid is generated at the app's start-up and remains constant throughout its lifetime. The app then uses the new waypoint's location to provide the user with guidance instructions. The policy actions, and waypoints by extension, are given relative to the current camera's pan-tilt orientation. The grid spans \SI{120}{\degree} in both the pan and tilt dimensions, giving a resolution of \SI{20}{\degree} per grid cell.

The position of the waypoint is given to the user by a set of guidance instructions (the arrows seen on-screen in Figure~\ref{fig:system-screenshot}) in terms of the actions defined in $\textbf{A}$. This might seem like an unnecessary extra step, but this system is designed for VI users who are limited to audio instructions, which makes it necessary to split the instructions up into separate channels that can be communicated more simply. This requirement becomes clear when one considers the bandwidth limitations of a human's audio channel as compared to the visual one.

By recording the previous positions and waypoint locations the app generated, it can infer the state values for $v$ and $n$. Furthermore, the camera provides the ID of the object in view, populating $o$ in the state. A real-time object detector was not used in this system to avoid unnecessary complexity and processing overhead and encoded QR codes and barcode scanner library were used to simulate object detection instead.

The QR code scanner requires raw image data. To avoid scanning multiple QR codes and using excess processing power by scanning large images, we only used the centre part of the screen to scan for a barcode. This scanner area is $300\times300$ pixels and scans for barcodes without causing any frame-rate drops.

@@

\noindent To test the effectiveness of our MDP and its policies, we designed a set of experiments that determined how effective the system is at directing a participant to point the device's camera towards a target object. This section discusses the experiments and their results.

\subsection{Experiment Design}

\noindent For the experiment, the MDP policies were integrated into an Android application that uses the camera to provide observation data and track the pose and viewing direction. We avoided adding unnecessary variability and complexity to the experiment by providing visual waypoints on screen which the participants were allowed to use. A generic environment mimicking a typical office desk layout was set up as the experiment environment and contains 7 different objects (i.e.\ encoded QR codes), one of which could be selected for each experiment run. See Figure~\ref{fig:env-picture} for a picture of the environment.

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{figures/test_env_picture.jpg}
  \caption{A picture of the environment used for the experiments. Each QR code represents an object.}\label{fig:env-picture}
\end{figure}

For each experiment, the participant was placed roughly \SI{1}{\meter} from the barcodes and was asked not to move from that spot during the experiment. The participant then pressed a button on the app that randomly selected a target object and was then guided towards that target by the app. Since the participants were allowed to use their vision, the targets were randomly selected and the participants were uninformed of what the target object was until it was found. This was done to avoid the participants learning the target objects' locations between each search run.

During in-house tests, we noted that the system sometimes makes guidance mistakes that lead the user towards an uncluttered edge of the search space, at which point the system had difficulty guiding the user back to the centre. To avoid this having an adverse impact and drawing out the experiments too long, we set a search step limit of 15, which means that the search was terminated when the number of waypoints exceeded 15. A search run therefore ended when the participant either successfully found the target object by pointing the phone camera to it and scanning the barcode, or exceeded the waypoint limit. After this, the participant then resets to the central position, selects a new target object and resumes the experiment.

The participants were left to repeat this process until we recorded 10 searches per object, giving us a total number of 70 recorded search samples per participant. We tested a total of 10 different participants (sex: 8 M, 2 F;\@ age: $\mu$ 31.2 years, $\sigma$ 8 years), most of which were sourced from our laboratory with none having any disabilities or handicaps. This gives our dataset a total of 700 search samples.

@@

\noindent We identified 3 different measures to evaluate the system's performance: target acquisition rate (TAR), number of steps to the target and the total time it took to find a target object. We present and discuss the results for each of these parameters in the next sections. The results for each individual participant is presented in Table~\ref{tab:results-full} and Table~\ref{tab:results-summary} gives the results across all of the participants. To provide a baseline measure for the ideal case, we ran a number of simulations in an environment mimicking the experiment setup with an agent that perfectly executes the policy, i.e.\ $u=u^*$. The TAR and steps to target results for the simulation are included in Table~\ref{tab:results-summary}.

>>>>

\title{Active Object Search with a Mobile Device\\for People with Visual Impairments}

\author{
  \authorname{Jacobus C. Lock{\sup{1}}, Grzegorz Cielniak\sup{1}, Nicola Bellotto\sup{1}}
  \affiliation{\sup{1}Lincoln Centre for Autonomous Systems (L-CAS), University of Lincoln, Lincoln, UK}
  \email{\{jlock, gcielniak, nbellotto\}@lincoln.ac.uk}
}

\keywords{Active vision, object search, visual impairment, Markov Decision Process}

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments, and their cameras in particular can be useful for assisting with tasks, such as reading signs or searching for objects in unknown environments. Previous research has looked at ways to solve these problems by processing the camera's video feed, but very little work has been done in {\em actively} guiding the user towards specific points of interest, maximising the effectiveness of the underlying visual algorithms. In this paper, we propose a control algorithm based on a Markov Decision Process that uses a smartphone's camera to generate real-time instructions to guide a user towards a target object. The solution is part of a more general active vision application for people with visual impairments. An initial implementation of the system on a smartphone was experimentally evaluated with participants with healthy eyesight to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help people with visual impairments find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more independent lives. Modern improvements in mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem. The work presented here is part of a project to assist people with visual impairments to navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}, expanding upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

\begin{figure}
  \centering
  \includegraphics[width=0.85\columnwidth]{figures/system_use.png}
  \caption{The system in use during an experiment. }\label{fig:system-use}
\end{figure}

@@

\noindent Assistive technology for people living with visual impairments is a growing research area~\cite{Manduchi2012,khoo2016multimodal}. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.\ Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. These type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWiz system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since a person with visual impairments is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.

Previous researchers have implemented active search and perception strategies in robots and image classifiers~\cite{Bajcsy2017} in an attempt to optimise their classification and planning tasks, for example by exploiting the structured nature of human environments and object placements. 
%
Two research teams have recently implemented an active object search strategy into their image classifiers~\cite{caicedo2015active,gonzalez2015active}. Their approaches use different methods but conceptually similar models to generate windows of interest for visual classification. The size and locations of the windows within the image are generated using the spatial relationship between objects, taken from the SUNCG and PASCAL datasets~\cite{song2016ssc,Everingham10}, and are iteratively changed based on the output from the respective models. The advantage of their approaches is that fewer windows are generated and submitted to the classifier, resulting in lower object classification times while still keeping state-of-the-art results for accuracy.

@@

In summary, much research has been conducted on recognition of and guidance towards target objects, including active vision solutions for image classifiers and robotic systems. However, to our knowledge, no previous work has been done on {\em active} object search and guidance for humans, which would especially benefit people with visual impairments. In this paper, we implement such an active vision system with a human in the loop
% ~\cite{bellotto2013,lock2017portable}
that guides the user towards an out-of-view target object. Our system exploits prior knowledge of the objects spatial distribution within an indoor environment, learned from a dataset of real-world images, and the history of past object observations made during the search.
%With this, we hope to make the searching process less random and move towards a more natural and efficient object searching process to the benefit of the VI\@.

@@

\noindent The work presented in this paper is a fundamental step towards a more general project's goal to develop a stand-alone system that can guide a person with visual impairments to his/her destination with minimal user input or intervention. A complete system diagram is given is Figure~\ref{fig:control-loop}. This closed-loop system is conceptually similar to other classical control problems, where the difference between desired and actual state of a process is used to generate a control signal that changes the process itself.

In this case, the reference signal, $r$, is the object the user wishes to capture with the smartphone's camera. The goal of the control block, $K$, is to generate human interpretable instructions, $u$, to guide the user towards the target object. The process to be controlled involves a human, $H$, who interprets the instruction and executes a physical action, $u^*$, to actually manipulate the smartphone's camera, $P$. A new observation, $y$, with the camera is then fed back to the loop and the error signal, $e$, is updated accordingly.

Here we focus in particular on the implementation of the control module $K$. Two important points are considered in the design of the controller. Firstly, $K$ must be scenario-agnostic, meaning that objects could be placed in different places with unknown a~priori information. Secondly, since each person could interpret the instruction $u$ differently (i.e. different transformation block $H$), the controller must be robust enough to handle such incorrect interpretations. For example, one person might interpret and execute an `UP' instruction correctly (i.e. $u \backsimeq u^*$), while another might interpret it correctly, but execute the wrong action. This risk can be mitigated by the use of clear and simple instructions that helps $u^*$ be as close as possible to $u$. The implementation of this controller is discussed in detail next.

@@

\noindent An MDP produces a policy of optimal actions for an agent to take in any given pre-defined state. In this case, the agent is defined as the guidance system and the policy is used to generate the next waypoint on the search path towards the target object. We assume fully observable states and known state transitions probabilities. The MDP is represented by the 5-tuple

\begin{equation}
  (\mathbf{S}, \mathbf{A}, \mathbf{T}, \mathbf{R}, \gamma),
\end{equation}
where $\mathbf{S}$ is a set of possible agent' states, $\mathbf{A}$ is a set of possible actions the agent can take in any given state, $\mathbf{T}$ is a set of state transition probabilities from state $s$ to state $s'$, with ${s,s'~\in~\mathbf{S}}$, and $\mathbf{R}$ is the reward the agent receives for reaching state $s'$ after executing action $a \in \mathbf{A}$ in state $s$. The scalar $\gamma$ is a discount factor that prioritises immediate over long-term rewards and which affects the model's convergence rate~\cite{Russell2009}. Each of these elements are defined and discussed next.

@@

The policy produced by an MDP defines the action the agent will take when it finds itself in any given state. In this case, the action is the direction of the next waypoint relatively to the current device's pose. The possible actions are given by

@@

To illustrate an example of actions sequence, let us consider the scene in Figure~\ref{fig:route-example}, which contains a number of simple, distinct objects (red boxes). The MDP guides the user in pointing the camera to the target object (the mug at the bottom-left of the figure). It does this by inferring the current state, which depends on the object currently observed by the camera, and generating an action, or instruction, that leads the user to the target object.
%The states are sampled and the actions are generated in discretised steps as the user moves the camera's view in the instructed direction. %This discrete behaviour is similar to a waypoint navigation approach where the user goes from one waypoint to the next, until the user points the camera to target object. 
An action is considered completed, and therefore a new state reached, when the camera has rotated more than a predefined angle or a new object is detected.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/office_desk_example.png}
  \caption{Example of action policies generated by the MDP to guide the user in pointing the camera from a random starting object (e.g. monitor) to a target object (e.g. mug). }\label{fig:route-example}
\end{figure}

@@

\noindent The reward $\mathbf{R}$ is the immediate reward that the agent receives after transitioning from state $s$ to state $s'$.
% and is given by 
% 
% \begin{equation}
%   R\in r(s, s').
% \end{equation} 
% 
The goal of the agent is to maximise its cumulative reward and it is very important to fine-tune $\mathbf{R}$ correctly for producing an effective action policy. In order to encourage the agent to find the target object as fast as possible, a relatively large positive reward should be assigned for successfully reaching the goal state, and a negative one in any other case. These parameters must be finely balanced to ensure effective object search behaviour.

@@

\begin{equation}
  \begin{aligned}
    o\in \mathbf{O} = \{monitor, mouse, keyboard, window,\\mug, stationery, desk, nothing\}.
  \end{aligned}
\end{equation}

The spatial relationships between the objects in $\mathbf{O}$ are extracted from the OpenImage dataset~\cite{openimages}, which consists of 1.74M images containing 14.6M manually drawn and labelled bounding boxes around objects (see Figure~\ref{fig:openimage-example} for some examples). The dataset is primarily aimed toward object recognition researchers to benchmark their models. In our case though, the bounding boxes and object labels are used to extract the spatial relationships between the different objects in $\mathbf{O}$. Since the camera perspectives and absolute distances between the objects in the images are not given, we can only extract the relationships in the basic action terms specified in $\mathbf{A}$, e.g. we can only say that object 1 is above object 2, but not how far above. Our relatively simple action-space is therefore suitable for the limited dataset information.

\begin{figure}
  \centering
  \begin{subfigure}
    \centering
    \includegraphics[width=\columnwidth]{figures/desk_example.png}
  \end{subfigure}
  \vfill
  \begin{subfigure}
    \centering
    \includegraphics[width=\columnwidth]{figures/mug_example.png}
  \end{subfigure}
  \caption{Examples of images from the OpenImage dataset containing objects from our set $\mathbf{O}$~\cite{openimages}. }\label{fig:openimage-example}
\end{figure}

Figure~\ref{fig:obj-relationships} shows the spatial relationship between a subset of $\mathbf{O}$ (desk, keyboard and mouse). For example, when the agent is in state ${s=\langle o=mouse, n, v \rangle}$ and is searching for the object $o_{target} = keyboard$, there is a strong probability that the target object is on the mouse's $LEFT$. The MDP of course will consider all of the objects' spatial relationships when generating the optimal policy.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/object_relationships.png}
  \caption{Examples of the spatial relationships between the desk, keyboard and mouse objects. Each square corresponds to the probability of executing an action (top square for $UP$, left square for $LEFT$, etc.)}\label{fig:obj-relationships}
\end{figure}

@@

\begin{equation}
  \alpha = \exp\Big(\frac{-i}{10~s_{tot}}\Big) - 0.001,
\end{equation}
where $i$ is the episode index. The discount factor $\gamma$ is set to 0.95 to prioritise long-term rewards and guarantee convergence.

Our MDP has a relatively small state-action space. Therefore, a solution can be found within a reasonable amount of time. However, it should be noted that adjusting the angle interval, or adding more actions or objects, can easily lead to an intractable space size.

\subsubsection{Waypoints Generation}\label{sec:waypoints-generation}

The system uses a 6$\times$6 discretised radial grid to simplify the tracking and waypoint generation processes. The grid spans \SI{120}{\degree} in both the pan and tilt dimensions, giving a resolution of \SI{20}{\degree} per grid cell. A policy action is converted by the system into a new search waypoint centred on a cell of the radial grid, e.g.\ an `$UP$' action will generate a waypoint one grid cell above the camera's current orientation. Note that this cell is not part of the MDP' state and the radial grid is only used to discretise the pan-tilt movements of the camera and to guarantee a minimum angular variation between subsequent actions.

% The grid is generated at the system's start-up and remains constant throughout its lifetime.
The system uses the waypoint's location to provide the user with guidance instructions (i.e. $u$ in Figure~\ref{fig:control-loop}). The policy actions, and waypoints by extension, are relative to the current camera's pan-tilt orientation. The grid is also wrapped so, if the location of a waypoint exceeds the \SI{120}{\degree} limit, the same waypoint is moved to the opposite side of the grid, effectively limiting the search space to a \SI{120}{\degree}$\times$\SI{120}{\degree} area.

\subsubsection{Smartphone Application}\label{sec:smartphone-application}

We incorporated the trained system into an app (see Figure~\ref{fig:system-screenshot}) for an Asus ZenPhone AR smartphone, running Android 7.0, with Google's augmented reality toolkit (ARCore), which provides the necessary 3D pose of the device. No further software or hardware modifications were required. This app is responsible for generating the guidance instructions and tracking the camera sensor ($K$ and $P$ blocks in Figure~\ref{fig:control-loop}) throughout a search session. Tracking the camera's pose allows the app to infer the current state and choose the optimal action to take next.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/system_screenshot2.png}
  \caption{A screenshot of the smartphone interface showing an example of guidance instruction (down-left in this case) towards a waypoint and the QR-object scanner area.}\label{fig:system-screenshot}
\end{figure}

The system determines the state values for $n$ (number of search steps so far) and $v$ (waypoint already visited or not) described in Section~\ref{sec:states}, by recording the previous positions and waypoint locations. The camera provides the ID of the object currently within view, which is assigned to the state variable $o$. In the current implementation, we did not use a real object detector, but we simulated it with 7 different QR codes, one for each unique object, and a camera-based QR code scanner from Android's machine learning API (MLKit). This simplification guarantees full observability of the state and let us focus on the performance of the MDP-based controller in the following experiments. 
%
% The QR scanner requires raw image data.
Moreover, to speed up the image processing and avoid scanning multiple QR codes, we only used the central part of the camera's frame, which is $300\times300$ pixels. This choice also defines the precision required in pointing the camera towards the object (see Figure~\ref{fig:system-screenshot}).

In a real application for people with visual impairments, the position of the waypoint would be given to the user by a set of audio or vibrotactile instructions. However, since we are mainly interested in evaluating the control algorithm of our system and not the interface ($K$ and not $u$), our current prototype generates guidance instructions with four on-screen arrows (see Figure~\ref{fig:system-screenshot}).
% This might seem like an unnecessary extra step, but this system is designed for user with visual impairments who are limited to audio instructions, which makes it necessary to split the instructions up into separate channels that can be communicated more simply.
% This requirement becomes clear when one considers the bandwidth limitations of a person's audio channel compared to the visual one.
Obviously, this visual interface is only used for debugging and experimental evaluation of the controller, and it will be replaced by an opportune audio interface, e.g.~\cite{bellotto2013}, at a second stage.

@@

\noindent To evaluate our system, we designed a set of experiments that determined how effective the MDP and its policies are at guiding the user in an object search task with the smartphone's camera. Since the focus of this work is on the algorithm for active object search, and not on the actual human performance, in the following experiments the system was tested by participants without any significant visual impairment. As explained in the previous section, this simplified the experimental design, allowing us to use the smartphone's display for the guidance instructions.
% This section discusses the experiments and their results.

\subsection{Experimental Design}

\noindent For the experiment, the MDP policies were integrated into an Android application that uses the camera to provide observation data and track the pose and viewing direction. Guidance instructions towards the waypoints were visualised on the screen, which the participants were allowed to use. The experimental environment mimicked a typical office desk layout and contained 7 different objects (i.e.\ encoded QR codes), one of which could be selected for each experiment run. See Figure~\ref{fig:env-picture} for a snapshot of the environment.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/test_env_picture.jpg}
  \caption{A snapshot of the environment used for the experiments. Each QR code represents an object.}\label{fig:env-picture}
\end{figure}

For each experiment, the participant was placed approximately at 1m from the closest barcodes and was asked to remain on that spot during the experiment. The participant started by pressing a button on the app, which randomly selected a target object and then guided the user towards it. Since the participants were allowed to use the smartphone's display, the target was randomly selected by the app without informing them, at least until it was found. This prevented the participants from learning the target objects' locations between subsequent runs of the experiment.

To avoid pointing at uncluttered edges of the search space, where the system had difficulty guiding the user back to the centre, we set a search step-limit of 15, which means the search was terminated when the number of waypoints generated by the system exceeded 15. A search run therefore ended when the participant either successfully found the target object by pointing the phone camera to it and scanning the barcode, or exceeded the waypoint limit. After this, the participant then restarted from the central position, generating a new random target object and repeating the experiment.

We recorded 10 searches per object, giving us a total number of 70 search samples per participant. The system was tested by 10 participants, mostly from our research group. Our final dataset consisted therefore of 700 search samples, which are analysed next.

@@

\noindent We identified 3 different measures to evaluate the system's performance: target acquisition rate (TAR), number of steps to the target and the total time it took to find a target object. We present and discuss the results for each of these parameters in the next sections. The results for each individual participant are presented in Table~\ref{tab:results-full}, while those across all the participants are shown in Table~\ref{tab:results-summary}. To provide a baseline measure for the ideal case, we ran a number of simulations in an environment mimicking the experiment setup with a ``virtual'' user who perfectly executes the policy, i.e.\ $u=u^*$. The TAR and steps to target results for the simulation are included in Table~\ref{tab:results-summary}.
